---
title: "Week7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Prerequisite - tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
```
#We will use the mpg dataset to create plots
```{r message=FALSE, warning=FALSE}
mpg
```
##Creating a ggplot
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```
##Let us add color to the graph based on class of the car
```{r message=FALSE, warning=FALSE}
ggplot(data=mpg)+geom_point(aes(x=displ, y=hwy, color=class))
```
##we can also change the size of the datapoints using class
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```
##We can also change the transpareny of the points and shape of the points
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```
##let us split up the plots into separate plots by class using facets
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```
##Let us create a geom - A geom is the geometrical object that a plot uses to represent data.
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))


ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy)) 
```
##We can also make groups and add color to it
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )
```
##We can also display multiple geoms in the same plot by adding multiple geom functions to ggplot():
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```
##Let us add a smooth line for class subcompact
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```
##Let us create a bar chart using data diamonds
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```
##Lets add color to this barchart
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```
##Lets add layers using 'clarity'
```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```
##Let us look at the coordinate systems, and the flipped coordinate systems
```{r message=FALSE, warning=FALSE}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```
##coord_quickmap() allows us to look at maps and set the aspect ratio correctly
```{r message=FALSE, warning=FALSE}
library(maps)
nz <- map_data("nz")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```

##Coord_polar() uses polar coordinates
```{r message=FALSE, warning=FALSE}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```
