---
title: "Week9"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###Prerequisites
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(lattice)
library(datasets)
```
###Let us add a smooth line to view the trend on plots
```{r message=FALSE, warning=FALSE}
with(airquality, 
     {plot(Temp, Ozone)+lines(loess.smooth(Temp, Ozone))})
```

###Let us add an annotation to the plot
```{r message=FALSE, warning=FALSE}
with(cars, plot(speed, dist))
title("Speed vs. Stopping distance")
```

###Let us create a lattice plot. 
###They tend to be most useful for conditioning types of plots, i.e. looking at how a dependent variable changes with independent variable across levels of a third variable
```{r message=FALSE, warning=FALSE}
state <- data.frame(state.x77, region = state.region)
xyplot(Life.Exp ~ Income | region, data = state, layout = c(4, 1))
```

###Let us create a scattedr plot with qplot using data mpg
```{r message=FALSE, warning=FALSE}
qplot(displ, hwy, data = mpg)
```

###Chapter 8: Graphic Devices
###Let us create a graph using plot()
```{r message=FALSE, warning=FALSE}
with(faithful, plot(eruptions, waiting))
title(main = "Old Faithful Geyser data")  
```

###We can create plots using pdf files
```{r message=FALSE, warning=FALSE}
## Open a PDF file from directory
pdf(file = "myplot.pdf")  
## Create a plot that will be saved to a file
with(faithful, plot(eruptions, waiting))  
title(main = "Old Faithful Geyser data")
## Close the PDF file
while (!is.null(dev.list()))  dev.off()
```