---
title: "Week12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Checking assumptions
```{r message=FALSE, warning=FALSE}
sim_1 = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x + rnorm(n = sample_size, mean = 0, sd = 1)
  data.frame(x, y)
}

sim_2 = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x + rnorm(n = sample_size, mean = 0, sd = x)
  data.frame(x, y)
}

sim_3 = function(sample_size = 500) {
  x = runif(n = sample_size) * 5
  y = 3 + 5 * x ^ 2 + rnorm(n = sample_size, mean = 0, sd = 5)
  data.frame(x, y)
}
```

##lets look at fitted versus residual plot
```{r message=FALSE, warning=FALSE}
set.seed(42)
sim_data_1 = sim_1()
head(sim_data_1)
plot(y ~ x, data = sim_data_1, col = "grey", pch = 20,
     main = "Data from Model 1")
fit_1 = lm(y ~ x, data = sim_data_1)
abline(fit_1, col = "darkorange", lwd = 3)
```
##Model with non constant variance
```{r message=FALSE, warning=FALSE}
set.seed(42)
sim_data_2 = sim_2()
fit_2 = lm(y ~ x, data = sim_data_2)
plot(y ~ x, data = sim_data_2, col = "grey", pch = 20,
     main = "Data from Model 2")
abline(fit_2, col = "darkorange", lwd = 3)
```

##Quadratic model
```{r message=FALSE, warning=FALSE}
set.seed(42)
sim_data_3 = sim_3()
fit_3 = lm(y ~ x, data = sim_data_3)
plot(y ~ x, data = sim_data_3, col = "grey", pch = 20,
     main = "Data from Model 3")
abline(fit_3, col = "darkorange", lwd = 3)
```
##Breusch-Pagan Test
```{r message=FALSE, warning=FALSE}
library(lmtest)
bptest(fit_1)
bptest(fit_2)
bptest(fit_3)
```

##Lets create histograms to look at the data
```{r message=FALSE, warning=FALSE}
par(mfrow = c(1, 3))
hist(resid(fit_1),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_1",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)
hist(resid(fit_2),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_2",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)
hist(resid(fit_3),
     xlab   = "Residuals",
     main   = "Histogram of Residuals, fit_3",
     col    = "darkorange",
     border = "dodgerblue",
     breaks = 20)
```
##Lets look at qqplot of residuals
```{r message=FALSE, warning=FALSE}
qqnorm(resid(fit_1), main = "Normal Q-Q Plot, fit_1", col = "darkgrey")
qqline(resid(fit_1), col = "dodgerblue", lwd = 2)
```
##Shapiro Wilk test - done to review the effectiveness of the model
```{r message=FALSE, warning=FALSE}
set.seed(42)
shapiro.test(rnorm(25))
shapiro.test(rexp(25))
#let us look at the fits
shapiro.test(resid(fit_1))
shapiro.test(resid(fit_2))
shapiro.test(resid(fit_3))
```
##Unusual observations
```{r message=FALSE, warning=FALSE}
par(mfrow = c(1, 3))
set.seed(42)
ex_data  = data.frame(x = 1:10,
                      y = 10:1 + rnorm(n = 10))
ex_model = lm(y ~ x, data = ex_data)

# low leverage, large residual, small influence
point_1 = c(5.4, 11)
ex_data_1 = rbind(ex_data, point_1)
model_1 = lm(y ~ x, data = ex_data_1)
plot(y ~ x, data = ex_data_1, cex = 2, pch = 20, col = "grey",
     main = "Low Leverage, Large Residual, Small Influence")
points(x = point_1[1], y = point_1[2], pch = 1, cex = 4, col = "black", lwd = 2)
abline(ex_model, col = "dodgerblue", lwd = 2)
abline(model_1, lty = 2, col = "darkorange", lwd = 2)
legend("bottomleft", c("Original Data", "Added Point"),
       lty = c(1, 2), col = c("dodgerblue", "darkorange"))

# high leverage, small residual, small influence
point_2 = c(18, -5.7)
ex_data_2 = rbind(ex_data, point_2)
model_2 = lm(y ~ x, data = ex_data_2)
plot(y ~ x, data = ex_data_2, cex = 2, pch = 20, col = "grey",
     main = "High Leverage, Small Residual, Small Influence")
points(x = point_2[1], y = point_2[2], pch = 1, cex = 4, col = "black", lwd = 2)
abline(ex_model, col = "dodgerblue", lwd = 2)
abline(model_2, lty = 2, col = "darkorange", lwd = 2)
legend("bottomleft", c("Original Data", "Added Point"),
       lty = c(1, 2), col = c("dodgerblue", "darkorange"))

# high leverage, large residual, large influence
point_3 = c(14, 5.1)
ex_data_3 = rbind(ex_data, point_3)
model_3 = lm(y ~ x, data = ex_data_3)
plot(y ~ x, data = ex_data_3, cex = 2, pch = 20, col = "grey", ylim = c(-3, 12),
     main = "High Leverage, Large Residual, Large Influence")
points(x = point_3[1], y = point_3[2], pch = 1, cex = 4, col = "black", lwd = 2)
abline(ex_model, col = "dodgerblue", lwd = 2)
abline(model_3, lty = 2, col = "darkorange", lwd = 2)
legend("bottomleft", c("Original Data", "Added Point"),
       lty = c(1, 2), col = c("dodgerblue", "darkorange"))

#Slop of the line
coef(ex_model)[2]
coef(model_1)[2]
coef(model_2)[2]
coef(model_3)[2]
```
##Let us look at the leverage for this dataset
```{r message=FALSE, warning=FALSE}
lev_ex = data.frame(
  x1 = c(0, 11, 11, 7, 4, 10, 5, 8),
  x2 = c(1, 5, 4, 3, 1, 4, 4, 2),
  y  = c(11, 15, 13, 14, 0, 19, 16, 8))

plot(x2 ~ x1, data = lev_ex, cex = 2)
points(7, 3, pch = 20, col = "red", cex = 2)

X = cbind(rep(1, 8), lev_ex$x1, lev_ex$x2)
H = X %*% solve(t(X) %*% X) %*% t(X)
diag(H)

sum(diag(H))
```

##We can use hat values method alternatively
```{r message=FALSE, warning=FALSE}
lev_fit = lm(y ~ ., data = lev_ex)
hatvalues(lev_fit)
#coefficient
coef(lev_fit)
#highest leverage
which.max(hatvalues(lev_fit))
lev_ex[which.max(hatvalues(lev_fit)),]
```
##creating a copy of the data
```{r message=FALSE, warning=FALSE}
lev_ex_1 = lev_ex
lev_ex_1$y[1] = 20
lm(y ~ ., data = lev_ex_1)
```
```{r message=FALSE, warning=FALSE}
which.min(hatvalues(lev_fit))

lev_ex[which.min(hatvalues(lev_fit)),]

lev_ex_2 = lev_ex
lev_ex_2$y[4] = 30
lm(y ~ ., data = lev_ex_2)

```
##let us look at the hat values of the models
```{r message=FALSE, warning=FALSE}
hatvalues(model_1)
hatvalues(model_2)
hatvalues(model_3)
```

##looking for large values
```{r message=FALSE, warning=FALSE}
hatvalues(model_1) > 2 * mean(hatvalues(model_1))

hatvalues(model_2) > 2 * mean(hatvalues(model_2))

hatvalues(model_3) > 2 * mean(hatvalues(model_3))
```