title: "EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}

library(nycflights13)
library(tidyverse)

```

##using the dataset flights
```{r message=FALSE, warning=FALSE}
flights
```

##let us filter the dataset and look at data from January 1st.
```{r message=FALSE, warning=FALSE}
filter(flights, month == 1, day == 1)
jan1 <- filter(flights, month == 1, day == 1)
```

##The following code finds all flights that departed in November or December:
```{r message=FALSE, warning=FALSE}
filter(flights, month == 11 | month == 12)
```
##short for values of x in y
```{r message=FALSE, warning=FALSE}
nov_dec <- filter(flights, month %in% c(11, 12))
```

##check for flights that were not delayed for more than 2 hours
```{r message=FALSE, warning=FALSE}
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```
##Comparison for missing values
```{r message=FALSE, warning=FALSE}
NA > 5
10 == NA
NA + 10
NA / 2
```
##function for checking missing values
```{r message=FALSE, warning=FALSE}
x <- NA

y <- NA

x == y
is.na(x)
```

##function filter only returns values that are TRUE or not missing
```{r message=FALSE, warning=FALSE}
df <- tibble(x = c(1, NA, 3))
filter(df, x > 1)
```
##missing values have to be called separately
```{r message=FALSE, warning=FALSE}
filter(df, is.na(x) | x > 1)
```

##arrange() works similarly to filter() except that instead of selecting rows, it changes their order.  
```{r message=FALSE, warning=FALSE}
arrange(flights, year, month, day)
```
##Select columns with select()
```{r message=FALSE, warning=FALSE}
select(flights, year, month, day)

# Selecting all columns that lie between between year and day (inclusive)
select(flights, year:day)

# Select all columns except those that lie between year to day (inclusive)
select(flights, -(year:day))
```
##renaming columns using select
```{r message=FALSE, warning=FALSE}
rename(flights, tail_num = tailnum)
```
##Adding new columns using mutate
```{r message=FALSE, warning=FALSE}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```
##Arithmetic operators using mutate
```{r message=FALSE, warning=FALSE}
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```
##summarize - collapses a dataframe into a single frame, ususally used with groupby
```{r message=FALSE, warning=FALSE}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))
```
##combining multiple operations
```{r message=FALSE, warning=FALSE}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
delay <- filter(delay, count > 20, dest != "HNL")
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```
##alternative to the above
```{r message=FALSE, warning=FALSE}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

##using counts() is good, because then you know you are not making conclusions based on a small amount of data
```{r message=FALSE, warning=FALSE}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))

delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )

ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```
##creating a scatter plot
```{r message=FALSE, warning=FALSE}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  )

ggplot(data = delays, mapping = aes(x = n, y = delay)) + 
  geom_point(alpha = 1/10)
```
##let us filter out the small groups
```{r message=FALSE, warning=FALSE}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)
```
## Convert to a tibble so it prints nicely
```{r message=FALSE, warning=FALSE}

batting <- as_tibble(Lahman::Batting)

batters <- batting %>% 
  group_by(playerID) %>% 
  summarise(
    ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    ab = sum(AB, na.rm = TRUE)
  )

batters %>% 
  filter(ab > 100) %>% 
  ggplot(mapping = aes(x = ab, y = ba)) +
    geom_point() + 
    geom_smooth(se = FALSE)
```
##using summary functions
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay1 = mean(arr_delay),
    avg_delay2 = mean(arr_delay[arr_delay > 0]) # the average positive delay
  )
```
## Why is distance to some destinations more variable than to others?
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))
```
## Lets look at when the first and last flights leave each day
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )
```
##Let us add a weight variable to count
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  count(tailnum, wt = distance)
```
## Let us look at how many flights left before 5am? 
```{r message=FALSE, warning=FALSE}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(n_early = sum(dep_time < 500))
#Proportion of flights
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(hour_perc = mean(arr_delay > 60))
```
##Grouping by multiple variables
```{r message=FALSE, warning=FALSE}
daily <- group_by(flights, year, month, day)
(per_day   <- summarise(daily, flights = n()))
(per_month <- summarise(per_day, flights = sum(flights)))
(per_year  <- summarise(per_month, flights = sum(flights)))
```
##Using ungrouping - opposite of grouping()
```{r message=FALSE, warning=FALSE}
daily %>% 
  ungroup() %>%             
  summarise(flights = n())
```
##Grouping using mutate and filter
```{r message=FALSE, warning=FALSE}
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests
```
##Standardize to compute per group metrics
```{r message=FALSE, warning=FALSE}
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```